Relationships in the extremes and their influence on competition and coexistence

### 2022-07-10 paper/Paper.Rnw
reorganizing ATA paragraph as suggeseted by LMH

%Currently 150 of an allowed 150 words for Ecology Letters
\begin{abstract}
\noindent Species coexistence attracts wide 
interest in ecology. Modern coexistence
theory (MCT) identifies coexistence mechanisms,
one of which, the storage effect, hinges on relationships between fluctuations
in environmental and competitive pressures. However, such relationships
are typically measured using covariance, which does not account for 
the possibility that environment and competition may be more related to each other
when they are strong than when weak, or vice versa. 
Recent work showed that such `asymmetric tail associations' (ATAs) are common
between ecological variables, and are important for 
extinction risk, ecosystem stability, and other phenomena. 
We extend the MCT, decomposing the storage effect to show the influence of ATAs.
Analysis of a simple model
and an empirical example using diatoms
illustrate that ATA influences can be comparable in magnitude to other
mechanisms of coexistence, and that ATAs can make the 
difference between species coexistence and competitive exclusion. 
ATA influences are an important new mechanism of coexistence. 
\end{abstract}

@@

%Intro to ATAs
Every ecological discipline involves the study of relationships among biological and environmental variables,
but standard %statistical methods 
measures for %examining the 
measuring the strength and direction of the association between two variables provide only 
%a small portion of the overall information content of the
partial information about the nature of the 
association \citep{Ghosh_2020_AER,Ghosh_2020_Eco,Ghosh_2020_ecolevol,Ghosh_2021}. For instance, 
the three pairs of variables in Fig. \ref{fig:noise} all have the same
Pearson correlation, but nevertheless demonstrate very different types of association. Other metrics 
of association are also often used, in addition to correlations
(e.g., various forms of regression), but correlations are extremely common, and other
methods also reveal only a small portion of the information content of the relationship between variables \citep{Nelsen2006}.
The variables in Fig. \ref{fig:noise}a,c 
exhibit what we have previously referred to as ``asymmetric tail association'' [ATA; \cite{Ghosh_2020_AER}], i.e., 
they are more strongly related to each
other in the left (Fig. \ref{fig:noise}a, termed ``left-tail association'') or right (Fig. \ref{fig:noise}c,
termed ``right-tail association'') portions of the variables' distributions. Although ATAs 
and related ideas had previously been seldom 
considered in ecology (but see \citealp{deValpine_2014, Anderson_2019, Popovic_2019}), 
our recent work shows that ATAs are extremely common \citep{Ghosh_2020_AER}, and can influence 
many phenomena, including extinction risk \citep{Ghosh_2020_Eco}, Taylor's law \citep{Ghosh_2020_AER}, 
synchrony, and stability of ecosystem functioning \citep{Ghosh_2020_ecolevol,Ghosh_2021}.
Although ATAs are now known to be important, the implications of ATAs for competition and species coexistence 
have, to our knowledge, not previously been considered; doing so is the overarching mission of this study.

>>>>

%Currently 150 of an allowed 150 words for Ecology Letters
\begin{abstract}
\noindent Species coexistence attracts wide 
interest in ecology. Modern coexistence
theory (MCT) identifies coexistence mechanisms,
one of which, the storage effect, hinges on relationships between fluctuations
in environmental and competitive pressures. However, such relationships
are typically measured using covariance, which does not account for 
the possibility that environment and competition may be more related to each other
when they are strong than when weak, or vice versa. 
Recent work showed that such `asymmetric tail associations' (ATAs) are common
between ecological variables, and are important for 
extinction risk, ecosystem stability, and other phenomena. 
We extend MCT, decomposing the storage effect to show the influence of ATAs.
Analysis of a simple model
and an empirical example using diatoms
illustrate that ATA influences can be comparable in magnitude to other
mechanisms of coexistence, and that ATAs can make the 
difference between species coexistence and competitive exclusion. 
ATA influences are an important new mechanism of coexistence. 
\end{abstract}

@@

%Intro to ATAs
Every ecological discipline involves the study of relationships among biological and environmental variables,
but standard %statistical methods 
measures for %examining the 
measuring the strength and direction of the association between two variables provide only 
%a small portion of the overall information content of the
partial information about the nature of the 
association \citep{Ghosh_2020_AER,Ghosh_2020_Eco,Ghosh_2020_ecolevol,Ghosh_2021}. 
For instance, consider the three pairs of variables in Fig. \ref{fig:noise}.
The variables in Fig. \ref{fig:noise}a,c exhibit what we have previously referred to as ``asymmetric tail association'' 
[ATA; \cite{Ghosh_2020_AER}], i.e., they are more strongly related to each
other in the left (Fig. \ref{fig:noise}a, termed ``left-tail association'') or right (Fig. \ref{fig:noise}c,
termed ``right-tail association'') portions of the variables' distributions. 
The relationships displayed in Fig. \ref{fig:noise} demonstrate very different types of association, 
but nevertheless all have the same Pearson correlation. 
Other metrics of association are also often used, in addition to correlations 
(e.g., various forms of regression), but correlations are extremely common, and other
methods also reveal only a small portion of the information content of the relationship between variables \citep{Nelsen2006}.
Although ATAs and related ideas had previously been seldom 
considered in ecology (but see \citealp{deValpine_2014, Anderson_2019, Popovic_2019}), 
our recent work shows that ATAs are extremely common \citep{Ghosh_2020_AER}, and can influence 
many phenomena, including extinction risk \citep{Ghosh_2020_Eco}, Taylor's law \citep{Ghosh_2020_AER}, 
synchrony, and stability of ecosystem functioning \citep{Ghosh_2020_ecolevol,Ghosh_2021}.
Although ATAs are now known to be important, the implications of ATAs for competition and species coexistence 
have, to our knowledge, not previously been considered; doing so is the overarching mission of this study.
